As was noted in chapter~\ref{chp:introduction}, measurements of the \acs{rpnd}
been mostly limited to the afterglow plasma or time-integrated quantites.
Electric field measurements, either with capacitive probes or nonlinear
wave-mixing, thus far provide the only detail of the \acs{rpnd} during its
development. Though the electric field can be used to estimate electron
densities and reaction rates in the plasma, this requires a number of additional
assumptions.

As a result, there is a lack of reliable information on the particle properties
of the \acs{rpnd} during its development. That said, such information is
essential to confirming the present understanding of how these discharges
develop, optimizing them for specific applications, and providing important
benchmarks for numerical simulations. Therefore, a clear need exists for direct
measurements of the \acs{rpnd} particle properties.

Unfortunately, these measurements present a significant challenge for most
traditional plasma diagnostics. In most situations, the obvious choice would be
the Langmuir probe for its simplicity and ease of implementation. However, the
fast variations in the plasma potential, slow response of the ion, and high
collisionality all preclude this approach \cite{Lieberman2005}. Furthermore, any
physical probe would act as a significant perturbation in the system.

The logical alternative to physical probes is the use of optical diagnostics,
however these have their own associated difficulties. Electrons cannot be
studied by their emissions because, with the exception of bremsstrahlung, they
do not emit light. This leaves the light emitted from excited atoms. Atomic
emission spectroscopy can be used to measure many different plasma quantities,
from electron density, to local electric field strength. Unfortunately,
spontaneous emission can be a slow process compared to the development of the
\acs{rpnd}. For example, the fastest neutral helium transition in visible
wavelengths (3$^3$D$_3$-2$^3$P$_2^\cdot$) has a radiative life of 14 ns
\cite{Ralchenko2011}.

This suggests that instead of waiting for spontaneous emission to occur, it may
be better to use some form of active spectroscopy. Though the added complexity
of a well-characterized light source is undesirable, it adds several interesting
possibilities. For example, Thomson scattering provides a means for direct
interaction with electrons. In addition, it has a high spatial and temporal
resolution and is able to measure the electron density and temperature
simultaneously \cite{VanGessel2012}. However, the electron density limit of
$5\times10^{12}$ cm$^{-3}$ is too low for use with the \acs{rpnd} which may have
electron densities well below this \cite{Pai2009}.

With the ability to directly interact with electrons, the next alternative is to
target one of the excited states of helium. The lowest such state, the triplet
metastable (2$^3$S), resides at 19.82 eV. This is a relatively large energy gap
for an atom and indicates that virtually no such states should exist at room
temperature. The triplet metastable (and all higher-energy states in helium)
will be populated exclusively by energetic electrons. Therefore, a measurement
of the triplet metastable level is a useful indicator of the degree of helium
excitation as the \acs{rpnd} develops, and could potentially be used to infer
some properties of the electron population.

Perhaps the most straightforward approach to a measurement of the triplet
metastable density is with absorption spectroscopy. This approach has a long
history in the study of metastables in gas discharges, going back at least six
decades \cite{Phelps1953}. At its most basic, the technique involves
illuminating a plasma with light matching a transition between the metastable,
and some upper level and measuring the amount of light transmitted. The amount
of light transmitted is proportional to the metastable density integrated along
the path of the light traversing the plasma. The bandwidth of this measurement
is only limited by the time required for the light to pass through the plasma.

\section{Setup}

Traditionally, the light used in absorption spectroscopy has been supplied by
discharge tubes with the same gas as the system under study. Though
straightforward, this approach is limited by the luminosity of the discharge
tube, and the fact that the emitted radiation is isotropic. More recently,
Millard et al. noted that diode lasers provide a greatly improved light source
that provides simple spatial selectivity, at intensities which can easily exceed
the saturation limit \cite{Millard1998}.

As with the study by Millard et al., the decision was made to study the
transition from the triplet metastable to the 2$^3$P$^\cdot_{0,1,2}$ (at
approximately 1083 nm). This was done for several reasons. For one, the closest
helium transition is over 7 nm away, making it relatively isolated. In addition,
the different levels or values of $J$ are all within the tuning range of a
single diode. As each level has a different degeneracy, $g$, the strength of
absorption varies depending on the selected level. Thus, the absorption strength
can be increased for low densities, or decreased at high densities, improving
the dynamic range of the diagnostic.

The laser used was a distribute feedback laser diode, produced by Toptica,
model LD-1083-0070-DFB-1. The specified linewidth of the laser was 3 MHz, well
below the natural linewidth of the transition, 10.2 MHz. This situation can be
exploited to directly measure the gas temperature of the system, as will be seen
in the next section. 


\section{Absorption Analysis}


\section{Results}


